Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-08-05 volunteer-clouds-ieee-ci-si.tex
Revisando comentarios
De @pacastillo, @mariosky y @fergunet. En general, si queréis reescribir una frase hacedlo y si acaso añadís un comentario diciendo por qué.

\begin{abstract}
Volunteer computing is used routinely in science and industry, but it
still has a long way to go if it is going to become mainstream in the
same way the cloud technologies in general have. The lack of general purpose tools
with analytics built in could be one of the reasons why this is so, as
well as the lack of proofs of concept. That is why in this paper we
present NodIO, a client-server architecture for distributed
evolutionary algorithms, written entirely in JavaScript, which needs
minimal server support to work and is able to support persistent,
asynchronous, distributed evolutionary algorithms running in the
browser. The architecture, along with some initial results on the expected performance and scaling
behavior is presented. We also use the algorithm analytics to try and  %[Pedro] I don't understand what "the algorithm analytics" stand for???
create a model of user behavior which, as is usual in this kind of
computing models, follow a Weibull distribution. 
\end{abstract}

@@

There is, then, space for systems that do not have any of those %[Mario] these assumptions? 
assumptions, such as the one presented in this paper, which will be
provisionally named {\sf NodIO}. NodIO is first an application that
runs in two tiers, a server and a client, the latter  can be possibly
run on a  % [Pedro] Both the server and the client in browsers?
          % different browsers as I understand? [JJ]  No, server in
          % server, client in browser, although there's nothing that
          % can't be done in a browser. 
browser. As such, it can be run in any computer that loads the web
page where the client is embedded, including possibly many over which%[Mario]Quitar el possibly?   
the scientist has no control.

Some time ago, our group faced the problem of diminishing funds for
buying new     % [Pedro] This sentence is so so long. I would split it
               % in two sentences. [JJ] Done.
hardware and continuous hardware failures of existing
clusters. 
%[Mario] Otra opción: Some time ago, our group faced the problem of diminishing funds 
%for buying new hardware. This was aggravated by the increasing maintance costs 
%resulting from the continuous failures of existing clusters.
Considering this, we leveraged our experience of working
with web applications, including JavaScript, since their inception and other volunteer and
unconventional distributed evolutionary computing systems, and we 
decided to designd and relase a new free framework that would allow anyone to
create a volunteer distributed EC experiment using cloud resources as
servers and browsers as clients.

@@

There have been relatively few efforts to analyze the
performance that can be obtained from these volunteer computing
experiments. 
%Too many efforts?
% No, too little - JJ 
% Digo efforts en la oración anterior :) - Mario
% Visto :-) Gracias- JJ
There were some initial attempts to simply dodge the issue by making
the algorithm adaptive to the kind of resources allotted to it
\cite{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can easily be
parallelized via population splitting or farming out evaluations to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms
\cite{gonzalez2010characterizing} which can, of course, be studied in
a more general distributed computing context
\cite{nogueras2015studying} or including it in a more general study of the
performance of the EA itself
\cite{DBLP:journals/gpem/LaredoBGVAGF14}.

%Mario: Estaba pensando que otra justificación muy fuerte para el uso de JS sobre plug-ins 
%ha sido la seguridad ya que estos presentan otra fuente de vulnerabilidad. Para complicar las cosas a veces (Flash, ActiveX) 
%son tecnologías cerradas y propietarias. Por lo menos creo que esas fueron las razones que se dieron para eliminar a Flash del iPhone.

@@

% [Pedro] desde el abstract no se había nombrado "NodIO". 
% Creo que en este párrafo debería introducirse, principalmente el nombre.
% Añado una frase presentándolo:
% [Mario] Nota: se menciona también en el segundo párrafo de la Introducción.
In this sense, in this paper we propose {\sf NodIO}, a cloud or metal based 
volunteer computing system, whose architecture
has been developed using JavaScript on both the client and server sides.
All parts of the system are free and available with a free
license from \url{https://github.com/JJ/splash-volunteer}.
Thus, {\sf NodIO} has two parts:\begin{enumerate}
\item A REST server, that is, a server that includes several remote %remote requests suena raro, propongo algo como: remote procedures o web interfaces   
  requests for storing and retrieving information (the CRUD cycle:
  create, request, update and delete) from the server. The information
  is of two kinds: {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
    GET}ing a random chromosome from it. It also peforms logging
  duties, but is basically a very lightweight and hign performance
  data store \cite{jj:idc:lowcost}. 
  %The information is of two kinds: Problem Based y ¿Cual otra? ¿Logging? No se entiende bien. ¿Podría ser problem domain?   
  The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found. 
\item A client that includes the evolutionary algorithm as 
  JavaScript code embedded in a web page that displays graphs and some
  additional links and information on the experiment. This code runs
  an evolutionary algorithm {\em island} starting with random
  population and sending, every 100 generations, the best individual
  back to the server (via a {\tt PUT} request) and requesting a random
  individual from the server (via a {\tt GET} request). Apart from
  these constants, the algorithm can be run in many different ways:
  stopping when solution is found or, as we have done in this paper,
  using WebWorkers, which run in the background even if the webpage or
  browser tab with the experiment is not currently on focus.  
\end{enumerate}

@@

The researcher only has to change a function to solve a different
problem, in this case the classical Trap function \cite{Ackley1987} has been 
used. JavaScript is a functional language and declaring a different
function and handling at the creation of the algorithm object, called %FERGU: poner algo tras el handling: handling it?
{\tt Classic} is the only change needed to work with a different
problem. Let us see how this system addresses the different challenges
outlined in the State of the Art (Section \ref{sec:soa}):\begin{itemize}
\item {\em Scalability} is provided via the use of a lightweight and
  high-performance, single-threaded, server based in NodeJS and
  Express.js. Although this single server is a bottleneck since it
  will eventually saturate, the fact that it runs as a single thread
  allows the service of many requests. In fact, a limit with the
  number of simultaneous requests will be reached, but so far it has
  not been found, unlike what we found in our previous systems, DCoR,
  \cite{gecco07:workshop:dcor}, which had a low scalability.
\item The system is {\em heterogeneous} since it does not need any
  performance, operating system or even browser requirement: anyone
  visiting the page, even from mobile devices, can load the algorithm. %Fergu: puesto lo de mobile devices para venderlo aún mejor
\item {\em Fault tolerance} is always an issue, and in this case, the
  single point of failure would be the server: the system, as such,
  would break down if the server fails. However, the individual
  islands in every browser would continue running, and having access
  to just one of them would allow the local algorithm to proceed. In
  fact, the island does not need the server to run: it runs locally if
  needed, with the only exception that it is obviously unable to
  communicate with the rest of the islands.
\item {\em Adaptiveness} is achieved simply through autonomous
  operation of every individual island and no synchronization
  mechanism. The islands in the system are, in fact, unaware of each
  other, communicating only through the server.
\item Since the algorithm is run on the browser, {\em safety} is
  achieved through its sandbox mechanisms. The user is thus assured
  that there is no unsafe access either to their local files or even
  to more resources that the browser should be allotted.
\item Running the algorithm is just a matter of loading the page,
  which means that operation is totally {\em anonymous}. For the same
  reason, {\em ease of use} is optimal, being as easy as simply
  clicking on an URL and available to anyone with access to a browser.
\item {\em Reasonable performance} is not ensured. In fact, we should
  make sure that there is a reasonable amount of clients over which
  the performance achieved is bigger than what you would obtain in
  your own desktop system. If that is not done, it is a pointless
  academic exercise.
\end{itemize}

>>>>

\begin{abstract}
Volunteer computing is used routinely in science and industry, but it
still has a long way to go if it is going to become a mainstream technology in the
same way the cloud technologies in general have. The lack of general purpose and open source tools
with analytics built in could be one of the reasons why this is so, as
well as the lack of proofs of concept. That is why in this paper we
present NodIO, a client-server architecture for distributed
evolutionary algorithms, written entirely in JavaScript, which needs
minimal server support to work and is able to support persistent,
asynchronous, distributed evolutionary algorithms running in the
browser. The architecture, along with some initial results on the expected performance and scaling
behavior is presented. We also use the algorithm data to  %[Pedro] I don't understand what "the algorithm analytics" stand for???
% data - JJ
find out what kind of patterns the user behaviour follow and try to fit them to some statistical distribution.
\end{abstract}

@@

There is, then, space for systems that do not have any of these %[Mario] these assumptions? [JJ] OK
assumptions, such as the one presented in this paper, which will be
provisionally named {\sf NodIO}. NodIO is first an application that
runs in two tiers, a server and a client, the latter  can be possibly
run on a  % [Pedro] Both the server and the client in browsers?
          % different browsers as I understand? [JJ]  No, server in
          % server, client in browser, although there's nothing that
          % can't be done in a browser. 
browser. As such, it can be run in any computer that loads the web
page where the client is embedded, including many over which%[Mario]Quitar el possibly?   [JJ] Done
the scientist has no control.

Some time ago, our group faced the problem of diminishing funds for
buying new     % [Pedro] This sentence is so so long. I would split it
               % in two sentences. [JJ] Done.
hardware. This was aggravated by the increasing maintance costs and extended downtime resulting from the continuous failures of existing clusters.
%[Mario] Otra opción: Some time ago, our group faced the problem of diminishing funds 
%for buying new hardware. This was aggravated by the increasing maintance costs 
%resulting from the continuous failures of existing clusters.
% [JJ] Done 
Considering this, we leveraged our experience in the design of web
applications, including JavaScript, since their inception and other
volunteer and 
unconventional distributed evolutionary computing systems, and we 
decided to designd and relase a new free framework that would allow anyone to
create a volunteer distributed EC experiment using cloud resources as
servers and browsers as clients.

@@

There have been relatively few efforts to analyze the
performance that can be obtained from these volunteer computing
experiments. 
There were some initial attempts to simply dodge the issue by making
the algorithm adaptive to the kind of resources allotted to it
\cite{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can easily be
parallelized via population splitting or farming out evaluations to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms
\cite{gonzalez2010characterizing} which can, of course, be studied in
a more general distributed computing context
\cite{nogueras2015studying} or including it in a more general study of the
performance of the EA itself
\cite{DBLP:journals/gpem/LaredoBGVAGF14}.

%Mario: Estaba pensando que otra justificación muy fuerte para el uso de JS sobre plug-ins 
%ha sido la seguridad ya que estos presentan otra fuente de vulnerabilidad. Para complicar las cosas a veces (Flash, ActiveX) 
%son tecnologías cerradas y propietarias. Por lo menos creo que esas
%fueron las razones que se dieron para eliminar a Flash del iPhone.
% [JJ] Escribe algo...

@@

% [Pedro] desde el abstract no se había nombrado "NodIO". 
% Creo que en este párrafo debería introducirse, principalmente el nombre.
% Añado una frase presentándolo:
% [Mario] Nota: se menciona también en el segundo párrafo de la
% Introducción.
% OK [JJ]
In this sense, in this paper we propose {\sf NodIO}, a cloud or metal based 
volunteer computing system, whose architecture
has been developed using JavaScript on both the client and server sides.
All parts of the system are free and available with a free
license from \url{https://github.com/JJ/splash-volunteer}.
Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
\item A REST server, that is, a server that includes several remote %remote requests suena raro, propongo algo como: remote procedures o web interfaces   
  requests for storing and retrieving information (the CRUD cycle:
  create, request, update and delete) from the server. The information
  is of two kinds: {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
    GET}ing a random chromosome from it, and {\em information} related
  to performance and state of the experiments. It also peforms logging
  duties, but is basically a very lightweight and hign performance
  data store \cite{jj:idc:lowcost}. 
  %The information is of two kinds: Problem Based y ¿Cual otra?
  %¿Logging? No se entiende bien. ¿Podría ser problem domain?   
% info. Añadido [JJ]
  The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found. 
\item A client that includes the evolutionary algorithm as 
  JavaScript code embedded in a web page that displays graphs and some
  additional links and information on the experiment. This code runs
  an evolutionary algorithm {\em island} starting with random
  population and sending, every 100 generations, the best individual
  back to the server (via a {\tt PUT} request) and requesting a random
  individual from the server (via a {\tt GET} request). Apart from
  these constants, the algorithm can be run in many different ways:
  stopping when solution is found or, as we have done in this paper,
  using WebWorkers, which run in the background even if the webpage or
  browser tab with the experiment is not currently on focus.  
\end{enumerate}

@@

The researcher only has to change a function to solve a different
problem, in this case the classical Trap function \cite{Ackley1987} has been 
used. JavaScript is a functional language and declaring a different
function and handing at the creation of the algorithm object, called
%FERGU: poner algo tras el handling: handling it?
%handing - JJ
{\tt Classic} is the only change needed to work with a different
problem. Let us see how this system addresses the different challenges
outlined in the State of the Art (Section \ref{sec:soa}):\begin{itemize}
\item {\em Scalability} is provided via the use of a lightweight and
  high-performance, single-threaded, server based in NodeJS and
  Express.js. Although this single server is a bottleneck since it
  will eventually saturate, the fact that it runs as a single thread
  allows the service of many requests. In fact, a limit with the
  number of simultaneous requests will be reached, but so far it has
  not been found, unlike what we found in our previous systems, DCoR,
  \cite{gecco07:workshop:dcor}, which had a low scalability.
\item The system is {\em heterogeneous} since it does not need any
  performance, operating system or even browser requirement: anyone
  visiting the page, even from mobile devices, can load the algorithm. %Fergu: puesto lo de mobile devices para venderlo aún mejor
\item {\em Fault tolerance} is always an issue, and in this case, the
  single point of failure would be the server: the system, as such,
  would break down if the server fails. However, the individual
  islands in every browser would continue running, and having access
  to just one of them would allow the local algorithm to proceed. In
  fact, the island does not need the server to run: it runs locally if
  needed, with the only exception that it is obviously unable to
  communicate with the rest of the islands.
\item {\em Adaptiveness} is achieved simply through autonomous
  operation of every individual island and no synchronization
  mechanism. The islands in the system are, in fact, unaware of each
  other, communicating only through the server.
\item Since the algorithm is run on the browser, {\em safety} is
  achieved through its sandbox mechanisms. The user is thus assured
  that there is no unsafe access either to their local files or even
  to more resources that the browser should be allotted.
\item Running the algorithm is just a matter of loading the page,
  which means that operation is totally {\em anonymous}. For the same
  reason, {\em ease of use} is optimal, being as easy as simply
  clicking on an URL and available to anyone with access to a browser.
\item {\em Reasonable performance} is not ensured. In fact, we should
  make sure that there is a reasonable amount of clients over which
  the performance achieved is bigger than what you would obtain in
  your own desktop system. If that is not done, it is a pointless
  academic exercise.
\end{itemize}
